A Switching Planner for Combined Task and Observation Planning

### 2010-11-22 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Only committing changes to yesterday's text.
 - Charles

In order to make assumptions available to the sequential planner, we
add one action $\actions(p_i;(T_i))$ to the problem for each element,
$p_i (T_i)$, of each probabilistic term from $(:init)$. The physics
of these actions is as follows. $\actions(p_i;(T_i))$ can be executed
if no $\actions(p_j;(T_j))$, $j \neq i$, has been executed from the
same probabilistic term, and, either
$(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct, or it
occurs in $T_k$ for some executed $\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^v$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^v$ is characterised by the union of
$s$ and atomic terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not necessarily true in the belief-state, we
must make two further adjustments to our sequential continual
planning. First, if $\action\in\actions$ has a precondition that is
not true with probability $1$ in the current state, then before it is
executed blindly, continual planning switches to a decision-theoretic
mode, that shall be discussed in detail below. Therefore, by
scheduling $\action$ the sequential planner makes a naive assumption
about the possibility of the agent eventually holding strong beliefs
about the truth values of specific state propositions. Second, if the
truth value of $\prop$ is not known, and no assumption is made about
it in the current abstract state, then if $\prop \in \poss(\action)$
we have that action $\action$ is not applicable.

Finally, we come to the issue of optimisation criteria. Here, we
propose the value of an $N$ horizon trace ---something we expect the
sequential planner to maximise--- corresponds to the maximal value of
that trace. In detail, where $p_i$ is the probability that the
$i^{th}$ execution is not a transition to $s^\bot$, we value a trace
according to:

@@

\noindent In goal directed problems, where the only non-zero reward is
received on the first transition to a goal state,
Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
infamous degenerate hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Moreover, form the perspective
of gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP, our optimal trace yields a maximal gradient
contribution under reasonable assumptions.

@@

>>>>

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\actions(p_i;(T_i))$ to the problem
for each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\actions(p_i;(T_i))$ can be executed if no
$\actions(p_j;(T_j))$, $j \neq i$, has been executed from the same
probabilistic term, and, either $(\pp{probabilistic}~..p_i~(T_i)..)$
is in the root conjunct, or it occurs in $T_k$ for some executed
$\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^v$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^v$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not uniformly true in the underlying
(resp. abstract) belief-state, we must make two further adjustments to
the problem posed to the sequential planner. First, if
$\action\in\actions$ has a precondition that is not true with
probability $1$, then before it is executed blindly, the sequential
continual planning switches to a decision-theoretic mode, discussed in
detail below. Overall, by scheduling $\action$ the sequential planner
makes a naive assumption about the possibility of the agent eventually
committing ---for the purpose of the current task--- to the truth
values of specific state propositions. Second, if the truth value of
$\prop$ is not known, and no assumption is made about it in the
current abstract state, then if $\prop \in \poss(\action)$ we have
that action $\action$ is not applicable.

%% In other words, the value of a proposition in an abstract
%% state is either {\em true}, {\em false}, or {\em unspecified} (i.e.,
%% \#). In the last case actions whose execution is preconditioned on the
%% term being truth functional are not available.

Finally, we come to the issue of valuing sequential plans, and
therefore giving the criterion that the sequential planner seeks to
optimise. Where $p_i$ is the probability that the $i^{th}$ action
execution is not a transition to $s^\bot$, we value a trace according
to:

@@

\noindent In goal directed problems, where the only non-zero reward is
received at the first transition to a goal state,
Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
infamous degenerate hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Form the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace yields a maximal
gradient step from a uniform soft-max policy.

@@

The sequential mode of our system always schedules decision-theoretic
planning before executing actions that are not applicable with
certainty. This is vastly inefficient if the utility of that plan is
not dependent on the actions successful execution. This is the case
where a plan is conformant. Therefore, in the future we must develop a
cheap (heuristic) procedure to evaluate the utility of sensing at
steps in sequential plan.
